### 2025-05-05 paper/ecai2025/sects/abstract.tex
few typos

\begin{abstract}
Reinforcement learning (RL) algorithms show promise in robotics and multi-agent systems but often suffer from low sample efficiency. While methods like SHAC leverage differentiable simulators to improve efficiency, they are limited to specific settings: they require fully differentiable environments (including transition and reward functions) and have primarily been demonstrated in single-agent scenarios. To overcome these limitations, we introduce \fname{}, a novel framework inspired by SHAC\@. \fname{} removes the need for differentiable simulator components by using neural networks to approximate the required gradients, training these networks alongside the standard policy and value networks. This allows the core SHAC approach to be applied in both non-differentiable and multi-agent environments. We evaluate \fname{} on challenging multi-agent tasks from the VMAS suite, comparing it against SHAC (where applicable) and PPO, a standard algorithm for non-differentiable settings. Our results demonstrate that \fname{} significantly outperforms PPO in both single- and multi-agent scenarios. Furthermore, in differentiable environments where SHAC operates, \fname{} achieves comparable performance despite lacking direct access to simulator gradients, thus successfully extending SHAC's benefits to a broader class of problems
\end{abstract}

>>>>

\begin{abstract}
Reinforcement learning (RL) algorithms show promise in robotics and multi-agent systems but often suffer from low sample efficiency. While methods like SHAC leverage differentiable simulators to improve efficiency, they are limited to specific settings: they require fully differentiable environments (including transition and reward functions) and have primarily been demonstrated in single-agent scenarios. To overcome these limitations, we introduce \fname{}, a novel framework inspired by SHAC\@. \fname{} removes the need for differentiable simulator components by using neural networks to approximate the required gradients, training these networks alongside the standard policy and value networks. This allows the core SHAC approach to be applied in both non-differentiable and multi-agent environments. We evaluate \fname{} on challenging multi-agent tasks from the VMAS suite, comparing it against SHAC (where applicable) and PPO, a standard algorithm for non-differentiable settings. Our results demonstrate that \fname{} significantly outperforms PPO in both single- and multi-agent scenarios. Furthermore, in differentiable environments where SHAC operates, \fname{} achieves comparable performance despite lacking direct access to simulator gradients, thus successfully extending SHAC's benefits to a broader class of problems.
\end{abstract}

### 2025-05-05 paper/ecai2025/sects/background.tex
few typos

In its trivial multi-agent extension, SHAC trains policies ${\pi_1}_\theta,\dots,{\pi_\na}_\theta$ by minimizing the loss function $\mathcal{L}_\theta^{\pi}:\trajectories^\bs\rightarrow\mathbb{R}$ for a batch of $\bs$ trajectories $\batch = \{\tau_1, \ldots, \tau_\bs\}$ sampled usin $\pi$. The loss $\mathcal{L}_\theta^{\pi}(\batch)$ is defined as:  
\begin{equation*}\label{eq:shac}
    \frac{1}{\bs\na\mh}\sum_{\tau\in\batch} \sum_{t=t_0}^{t_0+\mh-1} \sum_{i\in\agents} \discount^{t-t_0} R(s_{t}^\tau, a_{ti}^\tau) + \discount^\mh \valuef_\theta(s^\tau_{t_0+\mh})
\end{equation*}
where $\mh \ll \horizon$ is the rollout size (namely the number of steps used to estimate the return), $t_0$ is the initial time step of the rollout, $\valuef_\theta: \states \rightarrow \mathbb{R}$ is the value function estimating the expected return. It is apparent that SHAC assumes that the policy, reward, value, projection, and transition functions are all differentiable, i.e., $\{\policy{i}\}_{i\in\agents},\{\rewards{i}\}_{i\in\agents},\valuef_\theta, \text{and} \, \transition \in \mathcal{C}^1$. While this is a common requirement for policy and value functions, differentiable transition and reward functions are rarely available.

>>>>

In its trivial multi-agent extension, SHAC trains policies ${\pi_1}_\theta,\dots,{\pi_\na}_\theta$ by minimizing the loss function $\mathcal{L}_\theta^{\pi}:\trajectories^\bs\rightarrow\mathbb{R}$ for a batch of $\bs$ trajectories $\batch = \{\tau_1, \ldots, \tau_\bs\}$ sampled using $\pi$. The loss $\mathcal{L}_\theta^{\pi}(\batch)$ is defined as:  
\begin{equation*}\label{eq:shac}
    \frac{1}{\bs\na\mh}\sum_{\tau\in\batch} \sum_{t=t_0}^{t_0+\mh-1} \sum_{i\in\agents} \discount^{t-t_0} R(s_{t}^\tau, a_{ti}^\tau) + \discount^\mh \valuef_\theta(s^\tau_{t_0+\mh})
\end{equation*}
where $\mh \ll \horizon$ is the rollout size (namely the number of steps used to estimate the return), $t_0$ is the initial time step of the rollout, $\valuef_\theta: \states \rightarrow \mathbb{R}$ is the value function estimating the expected return. It is apparent that SHAC assumes that the policy, reward, value, projection, and transition functions are all differentiable, i.e., $\{\policy{i}\}_{i\in\agents},\{\rewards{i}\}_{i\in\agents},\valuef_\theta, \text{and} \, \transition \in \mathcal{C}^1$. While this is a common requirement for policy and value functions, differentiable transition and reward functions are rarely available.

### 2025-05-05 paper/ecai2025/sects/introduction.tex
few typos

Our contributions include:
\begin{compactitem}
    \item A novel RL framework (\fname{}) that employs learned gradient approximations, eliminating the need for differentiable simulators
    \item The first empirical evaluation of SHAC in multi-agent environments, establishing a baseline for differentiable MARL
    \item A comprehensive comparison across environments with varying agent counts and differentiability properties
    \item Experimental evidence demonstrating that \fname{} matches SHAC's performance in differentiable settings while substantially outperforming PPO in sample efficiency and final reward across both single-agent and complex multi-agent scenarios
\end{compactitem}

>>>>

\noindent Our contributions include:
\begin{compactitem}
    \item a novel RL framework (\fname{}) that employs learned gradient approximations, eliminating the need for differentiable simulators;
    \item the first empirical evaluation of SHAC in multi-agent environments, establishing a baseline for differentiable MARL;
    \item a comprehensive comparison across environments with varying agent counts and differentiability properties;
    \item experimental evidence demonstrating that \fname{} matches SHAC's performance in differentiable settings while substantially outperforming PPO in sample efficiency and final reward across both single-agent and complex multi-agent scenarios.
\end{compactitem}
